\addchap{Acknowledgements}
% \todo{thank people}
% \todo{more detail...}

Foremost, I would like to thank my supervisor \textbf{Ali Alavi} for his guidance and undying optimism, as well as for being an excellent professional role model. Similarly, I would like to thank \textbf{Daniel Kats} for his professional guidance and patience with me as I learned the basics of electronic structure theory, especially at the start of my PhD while I was still working remotely from Vancouver. I also extend a thanks to my committee members \textbf{Joris van Slageren} and \textbf{Andreas K\"ohn}.

Similarly, much of my professional development can be owed to mentors along the way, in particular \textbf{Darryl Barber}, \textbf{Christian Ast}, and \textbf{Roman Krems}.

I also thank \textbf{Evelin Christlmaier} for all her support, both professionally and personally, and with whom working has been a pleasure.While our time together in the group was short, \textbf{Werner Dobrautz} distilled in me a greater sense of mindfulness and reassurance that everything will be okay. Similarly, \textbf{Johannes Hauskrecht} has been a wonderful officemate to whom I coud always whine about life as a PhD student.

My time at MPI has, to my pleasure, involved a lot of software development, and for that I am especially grateful to have had such knowledgeable coworkers as \textbf{Robert Anderson} and \textbf{Oskar Weser} to go to for advice and with whom to argue which text editor is best.

I would also like to thank my former and present coworkers \textbf{Pablo L\'opez-R\'ios}, \textbf{Aron Cohen} and \textbf{Kai Guther} for their work on making \tchint a reality. \textbf{Andreea Filip}, \textbf{Kristoffer Simula} and \textbf{Ke Liao} have been especially pivotal in making the \pytchint package possible.


Navigating administration as a foreigner in Germany can be extremely difficult, but this stress has been lifted off my shoulders by \textbf{Birgit King}, to whom I am much in debt.

% \todo{Friends back home: Karan(!), Michael, Charlie, Zack, Cam, Mr Barber}

Finally, I want to thank my dear friends who made my time in Stuttgart infinitely more colourful: \textbf{Aybike Reyhanl\i}, \textbf{Thomas Schraivogel}, \textbf{Ingemar Schmidt}, \textbf{Rut Martinez}, \textbf{Nikos Papanikolaou}, \textbf{Martin Baumann}, and especially \textbf{Theresa Huber} with whom I shall always cherish our countless hours dancing together.
